<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{T Interval: One Sample}
-->
# T Interval

**Table of Contents**

1. T Statistic
    a. Definitions and Terminology
2. Confidence Limits
3. Study Design Considerations
4. Example: [Application]
    a. Sample Size
    b. Margin of Error
    c. Standard Deviation
    d. Significance
    e. One Sided Intervals
5. Study Design Derivations
    a. Sample Size
    b. Standard Deviation
    c. Significance
    d. One Sided Intervals
6. References

## 1. T Statistic
The $t$-statistic is a standardized measure of the magnitude of difference between a sample's mean and some known, non-random constant.  It is similar to a $z$-statistic, but differs in that a $t$-statistic may be calculated without knowledge of the population variance.

### 1a. Definitions and Terminology
Let $\theta$ be a sample parameter from a sample with standard deviation $s$.  Let $\theta_0$ be a constant, and $s_\theta = s/\sqrt{n}$ be the standard error of the parameter $\theta$.  $t$ is defined:
$$t = \frac{\theta - \theta_0}{s_\theta} = \frac{\theta - \theta_0}{\frac{s}{\sqrt{n}}}$$

## 2. Confidence Limits
The confidence interval for $\theta$ is written:
$$\theta \pm t_{1-\alpha/2} \cdot \frac{s}{\sqrt{n}}$$

The value of the expression on the right is often referred to as the _margin of error_, and we will refer to this value as 
$$E = t_{1-\alpha/2} \cdot \frac{s}{\sqrt{n}}$$

## 3. Study Design Considerations

## 4. Example: [Application]

### 4a. Sample Size
### 4b. Margin of Error
### 4c. Standard Deviation
### 4d. Significance
### 4e. One Sided Intervals

## 5. Study Design Derivations
### 5a. Sample Size
$$E = t_{1-\alpha/2} \cdot \frac{s}{\sqrt{n}}$$
$$\frac{E}{t_{1-\alpha/2}} = \frac{s}{\sqrt{n}}$$
$$\frac{E}{t_{1-\alpha/2} \cdot s} = \frac{1}{\sqrt{n}}$$
$$\frac{t_{1-\alpha/2} \cdot s}{E} = \sqrt{n}$$
$$\frac{t_{1-\alpha/2}^2 \cdot s^2}{E^2} = n$$

Since $t_{1-\alpha/2}$ depends on the value of $n$, this is not a problem that is easily reduced to a solution.  Many texts encourage using $z_{1-alpha/2}$ as a substitute, but we're using computers here, so we can probably do a little better.  Instead, if we write the last line as:
$$\frac{t_{1-\alpha/2}^2 \cdot s^2}{E^2} - n = 0$$
$$\big(\frac{t_{1-\alpha/2}^2 \cdot s^2}{E^2} - n\big)^2 = 0$$

We now have a quadratic equation.  We'll use the `optimize` function in the `stats` package to find a best solution for $n$.  

Consider when we have $n=25$, $s=4$ and $\alpha=.05$.  The value of $E$ here is 
$$E = t_{1-\alpha/2} \cdot \frac{s}{\sqrt{n}} = 2.063899 \cdot 4/5 = 1.651119$$.

Now let's rewrite the problem to solve for $n$ using `optimize`.
```{r}
fn <- function(n) (qt(1-.05/2, n-1)^2 * 4^2 / 1.651119^2 - n)^2
optimize(fn, c(0, 100))
```

On the other hand, using the $z$ approximation yields
```{r}
qnorm(1-.05/2)^2 * 4^2 / 1.651119^2
```

which is two subjects short of what we would actually need.  `n_t1samp_interval` uses the `optimize` function and searches over the values 0 to 1,000,000,000.

### 5b. Standard Deviation
$$E = t_{1-\alpha/2} \cdot \frac{s}{\sqrt{n}}$$
$$\frac{E}{t_{1-\alpha/2}} = \frac{s}{\sqrt{n}}$$
$$\frac{E \cdot \sqrt{n}}{t_{1-\alpha/2}} = s$$

### 5c. Significance
$$E = t_{1-\alpha/2} \cdot \frac{s}{\sqrt{n}}$$
$$\frac{E \cdot \sqrt{n}}{s} = t_{1-\alpha/2}$$
$$\Phi_t\Big(\frac{E \cdot \sqrt{n}}{s}\Big) = \Phi_t(t_{1-\alpha/2})$$
$$\Phi_t\Big(\frac{E \cdot \sqrt{n}}{s}\Big) = 1 - \frac{\alpha}{2}$$
$$1 - \cdot \Phi_t\Big(\frac{E \cdot \sqrt{n}}{s}\Big) = \frac{\alpha}{2}$$
$$2 \cdot \Big[1 - \Phi_t\Big(\frac{E \cdot \sqrt{n}}{s}\Big)\Big] = \alpha$$

## 6. References